---
title: "R Notebook"
output: html_notebook
---

Covid La

```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(viridis)
library(RColorBrewer)
```


```{r}
#covid <- read_csv("us-counties.csv")
covid <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
```

```{r}
covid
```

```{r}
covid %>% 
  filter(county == "Los Angeles") %>% 
  ggplot(aes(x = date, y = deaths)) + 
  geom_col() +
  cowplot::theme_minimal_vgrid(16) +
  theme(axis.title.y = element_blank(), 
        legend.position = "none") +
  xlab("number of covid-19 deaths in LA")
```
```{r}
covid %>% 
  filter(county == "Los Angeles") %>% 
  ggplot(aes(x = date, y = deaths)) + geom_line()
```

Here we'll plot the death rate on a line graph by hotspots. See there's a problem with New Orleans. This is because the parish New Orleans is in is "Orleans", but there's also an Orleans in New York, so the filtering below isn't working. Need to fix so we filter out Orleans Country New York. 

```{r}
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
covid %>% 
  filter(county %in% c("Orleans", "Los Angeles", "New York City", "King", "Santa Clara") & state %in% c("Louisiana", "California", "Washington", "New York")) %>% 
  ggplot(aes(x = date, y = deaths, color = county)) + geom_line() +
  scale_fill_brewer(palette = "YlOrRd") +
  theme(axis.title.x = element_blank(), 
        legend.position = "none") +
  cowplot::theme_minimal_vgrid(16) +
  facet_wrap(~county)
  #scale_color_viridis(discrete = TRUE)
  #scale_fill_manual(values = cbp1)
```

Let's calculate a rate of growth. Here let's use the death rate b/c testing is problematic as an indicator until capacity increases and more of the population is tested. 

```{r}
covid %>% 
  filter(county %in% c("Los Angeles")) %>% 
  arrange(date) %>% 
  mutate(diff_date = date - lag(date),  # Difference in time (just in case there are gaps)
         diff_growth = deaths - lag(deaths), # Difference in deaths between days
         rate_percent = (diff_growth / as.double(diff_date)) / deaths * 100) # growth rate in percent
```

